---
output: github_document
---

## fcod alternative assessment

* Drop age 1 in the input data
* Use default sam settings - at least in the base run

Running stuff:

* The code for the assessment run is found in R/fcod_base.R
* The results are in run:
  * fit.rds
  * retro.rds

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
# remotes::install_github("einarhjorleifsson/fishivce")
library(fishvice)

# The NWWG 2022 fall run (not to be confused with the ipb)
fit.nwwg2022 <- sam_get_fit("Cod_27_5b1")
rbx.nwwg2022 <- sam_rbx(fit.nwwg2022)
opr.nwwg2022 <-
  rbx.nwwg2022$opr |>
  mutate(fleet = case_when(str_sub(fleet, 1, 2) == "SP" ~ "spring",
                           str_sub(fleet, 1, 2) == "SU" ~ "summer",
                           TRUE ~ fleet),
         o = exp(o),
         p = exp(p))

# base run:
fit <- read_rds("run/fit.rds")
retro <- read_rds("run/retro.rds")
rbx <- sam_rbx(fit)
rby <- 
  rbx$rby |> 
  mutate(year = ifelse(variable == "rec", year - 1, year))
rbya <-
  rbx$rbya |>
  mutate(yc = year - age)
rbya.wide <- 
  rbya |> 
  select(-yc) |> 
  spread(variable, val)
opr <-
  rbx$opr |>
  mutate(fleet = case_when(str_sub(fleet, 1, 2) == "SP" ~ "spring",
                           str_sub(fleet, 1, 2) == "SU" ~ "summer",
                           TRUE ~ fleet),
         o = exp(o),
         p = exp(p))

# base retro:
retro <- read_rds("run/retro.rds")
assyears <- 2021 - 0:(length(retro) - 1)
rby.retro <- map(retro, sam_rby)
rbya.retro <- map(retro, sam_rbya)
opr.retro <- map(retro, sam_opr)
names(rby.retro)  <- names(rbya.retro) <- names(opr.retro) <- assyears

rby.retro <- 
  bind_rows(rby.retro, .id = "assyear") |> 
  mutate(assyear = as.integer(assyear)) |> 
  bind_rows(sam_rby(fit) |> mutate(assyear = 2022L))
```

## The problem

```{r}
opr.nwwg2022 |> 
  filter(fleet %in% c("spring", "summer")) |> 
  mutate(yc = year - age) |> 
  arrange(fleet, yc, age) |> 
  group_by(fleet, yc) |> 
  mutate(logU = log(o / lead(o))) |> 
  ungroup() |> 
  filter(age == 1) |> 
  ggplot(aes(year, logU)) +
  geom_point() +
  geom_smooth(span = 0.999) +
  facet_wrap(~ fleet, scales = "free_y") +
  labs(x = NULL, y = "Log catch ratio of age 1 vs 2")
```

## One solution

```{r}
rby.retro |> 
  filter(variable != "catch",
         year > 2000) |> 
  ggplot(aes(year, est, group = assyear)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_y")
rby.retro <- 
  rby.retro |> 
  left_join(sam_rby(fit) |> select(year, est2022 = est, variable)) |> 
  mutate(p = log(est / est2022)) |> 
  filter(variable != "catch",
         year > 2000)
bias <- 
  rby.retro |> 
  filter(between(year, 2005, 2018),
         year == assyear) |> 
  group_by(variable) |> 
  summarise(y1 = min(year),
            y2 = max(year),
            sd = sd(p),
            p = mean(p))
```


```{r fig.height = 8}
ggplot() +
  annotate("rect", xmin = 2000, xmax = 2023, ymin = -0.2, ymax = 0.2, fill = "pink") +
  geom_hline(yintercept = 0, colour = "white") +
  geom_segment(data = bias,
              aes(x = y1, xend = y2, y = p, yend = p),
              colour = "blue",
              lwd = 2) +
  geom_point(data = rby.retro |> filter(year == assyear),
             aes(year, p),
             colour = "red") +
  scale_x_continuous(limits = c(2000, 2023)) +
  geom_line(data = rby.retro,
            aes(year, p, group = assyear)) +

  facet_wrap(~ variable, scales = "free_y") +
  #scale_x_continuous(breaks = seq(2000, 2024, by = 2)) +
  scale_y_continuous(breaks = seq(-3, 3, by = 0.2)) +
  labs(x = NULL, y = "cv proxy")
```

## Caveats

* We still have some problems, but they are less than what we have seen at the NWWG 2022 fall meeting.
  * Should be able to ameleorate some of them with some addition tweeking

## Session info

```{r}
devtools::session_info()
```

